Associahedra for finite type cluster algebras \\ and minimal relations between $\b{g}$-vectors

### 2019-07-12 thirdPaper/draft/typeConeAssociahedra.tex
workout latex

% \section*{Main results and logical dependencies}
\subsection*{Structure of the paper: main results and logical dependencies}

@@

@@

\noindent {\bf Cluster algebras of finite type} (\cref{subsec:typeConeCA})

We first consider the~$\b{g}$-vector fan~$\Fan(\B_\circ)$ of a skew-symmetrizable cluster algebra of finite type, with respect to any initial exchange matrix of finite type~$\B_\circ$. We provide the following polytopal realizations of~$\Fan(\B_\circ)$:
% We consider the set~$\meshes$ of all pairs~$\{x,x'\}$ related by non-initial mesh mutations (\cref{def:meshMutation}), and we denote by~$\principalVariables$ the set of cluster variables.
% Let~$\{x,x'\}\in\meshes$, and let~$X,X'$ be two clusters such that~$X\setminus\{x\}=X'\setminus\{x'\}$.
% For any~$y\in\principalVariables$, define~$\alpha_{x,x',y}$ to be~$|b_{xy}|$ if~$y\in X\cap X'$ and 0 otherwise.
% By the unique exchange relation property, the~$\alpha_{x,x',y}$'s do not depend on the specific choice of the clusters~$X,X'$.

\noindent \cref{thm:allPolytopalRealizationsCA}: 
\emph{For any~$\b{\ell} \in \R_{>0}^{\meshes}$, the polytope
\[
R_{\b{\ell}}(\B_\circ) \eqdef \Bigset{\b{z} \in \R^{\principalVariables}}{\b{z} \ge 0 \text{ and } \b{z}_x + \b{z}_{x'} - \sum_{y \in \principalVariables} \alpha_{x,x',y} \, \b{z}_{y} = \b{\ell}_{\{x,x'\}} \text{ for all } \{x,x'\} \in \meshes}
\]
is a generalized associahedron, whose normal fan is the cluster fan~$\gvectorFan[\B_\circ]$.
Moreover, the polytopes~$R_\b{\ell}(\B_\circ)$ for~$\b{\ell} \in \R_{>0}^{\meshes}$ describe all polytopal realizations of~$\gvectorFan[\B_\circ]$.}

@@

Some crucial pieces of the proof of~\cref{thm:allPolytopalRealizationsCA} use cluster categories (2-Calabi--Yau triangulated categories) and are deferred to~\cref{sec:clusterCategories}. Namely, a reformulation of~\cite[Thm 7.5]{BuanMarshReinekeReitenTodorov}, proved in~\cref{sec:clusterCategories}, gives the first step towards applying our type cone strategy.

\noindent \cref{prop:uniqueExchangePropertyCA}: \emph{The cluster fan~$\Fan(\B_\circ)$ has the unique exchange relation property.}

@@

\noindent \cref{coro:simplicialTypeConeCA}: \emph{The type cone~$\typeCone(\Fan(\B_\circ))$ is simplicial.}

\medskip
\bigskip

\noindent {\bf Gentle algebras, non-kissing complex, generalized accordiohedra} (\cref{subsec:typeConeNKC})

The same strategy is applied to a class of finite dimensional algebras, called gentle algebras, whose representations are combinatorially well understood.
We identify two conditions, called brick and 2-acyclicity (\cref{def:brick2acyclic}), that enable us to make use of the type cone approach.
We note that the~$\tau$-tilting theory of gentle algebras is the algebraic counterpart of the combinatorics of accordions on surfaces~\cite{PaluPilaudPlamondon-nonkissing,BrustleDouvilleMousavandThomasYildirim,PaluPilaudPlamondon-surfaces}.

Let~$\quiver = (Q,I)$ be a gentle bound quiver, and let~$\Fan(\quiver)$ be the fan of its~$\b{g}$-vectors (\cref{subsubsec:nonkissingComplex}), also called its non-kissing fan.
We prove \cref{prop:exchangeablePairsNKC} (ii) in \cref{sect::Kbproj}, by making use of representation theory, and state here the following consequence:

\noindent \cref{corollary: uerp gentle}: \emph{The non-kissing fan~$\Fan(\quiver)$ has the unique exchange relation property.}
\arnau{In the proof of \cref{prop:exchangeablePairsNKC} it says that the UERP follows from PPP17}\pg{Not for point (ii), which uses results of Section 4}
\medskip

\cref{subsubsec:simplicialTypeConeNKC} is dedicated to giving a purely combinatorial proof of:

\noindent \cref{coro:simplicialTypeConeNKC}: \emph{Assume that the gentle bound quiver~$\quiver$ is brick and 2-acyclic.
Then the type cone~$\typeCone(\Fan(\quiver))$ is simplicial.}

An alternative algebraic proof, using the main result of \cref{sec:extricats}, is given in \cref{sect::Kbproj}.

Combining these pieces, we make use of our type cone approach to deduce:

\noindent \cref{coro:allPolytopalRealizationsNKC}: \emph{For any brick and $2$-acyclic gentle quiver~$\quiver$ and any~$\b{\ell} \in \R_{>0}^{\strings}$, the polytope
\[
R_\b{\ell}(\quiver) \eqdef \set{\b{z} \in \R^{\walks}}{\begin{array}{l} \b{z} \ge 0 \qquad\text{and}\qquad \b{z}_\omega = 0 \text{ for any improper walk } \omega \\ \b{z}_{\cc{\sigma}} + \b{z}_{\hh{\sigma}} - \b{z}_{\hc{\sigma}} - \b{z}_{\ch{\sigma}} = \b{\ell}_\sigma \text{ for all } \sigma \in \strings\end{array}}
\]
is a realization of the non-kissing fan~$\gvectorFan[\quiver]$.
Moreover, the polytopes~$R_\b{\ell}(\quiver)$ for~$\b{\ell} \in \R_{>0}^{\strings}$ describe all polytopal realizations of~$\gvectorFan[\quiver]$.}

@@

\bigskip

\noindent {\bf Graphical nested complexes} (\cref{subsec:typeConeGA})

@@

\noindent \cref{cor:typeconenestedfan}:\emph{
Consider a graph~$\graphG$ on~$\ground$ with tubes~$\tubes$. 
Then, for a height vector~$\b{h} \in \R^{\tubes}$, the nested fan~$\gvectorFan[\graphG]$ is the normal fan of the graph associahedron
\[
\set{\b{x} \in \R^\ground}{\dotprod{\gvector{\tube}}{\b{x}} \le \b{h}_{\tube} \text{ for any tube } \tube \in \tubes}
\]
if and only if~$\b{h}_{\varnothing} = \b{h}_{\graphG} = 0$ and
 $\b{h}_{\tube[s] \ssm \{v'\}} + \b{h}_{\tube[s] \ssm \{v\}} > \b{h}_{\tube[s]} + \b{h}_{\tube[s] \ssm \{v, v'\}}$ for any tube~$\tube[s] \in \tubes$ and distinct non-disconnecting vertices~$v,v'$ of~$\tube[s]$.
}

\bigskip

\noindent {\bf Cluster categories}

@@

Let~$\cat$ be a cluster category with finitely many isomorphism classes of indecomposable objects (see \cref{sect::setting} for the precise and more general setting in which the theorem is proven).
Let~$\Ksp(\cat)$ be the split Grothendieck group of~$\cat$.
Fix a cluster-tilting object~$T\in\cat$, and let~$K_0(\cat ; T)$ be the quotient of~$\Ksp(\cat)$ by the relations~$[X]+[Z]-[Y]$ for all triangles
$
X\xrightarrow{} Y \xrightarrow{} Z \xrightarrow{h} \susp X
$
with~$h \in (\susp T)$.
Denote by~$\b{g}:\Ksp(\cat) \to K_0(\cat ; T)$ the canonical projection.
For any indecomposable~$X\in\cat$, write~$\ell_X = [X]+[\susp^{-1}X]-[E]$, where~$E$ is the middle term of an almost split triangle starting at~$X$.
For any objects~$X,Y\in\cat$, we write~$\langle X, Y \rangle$ for~$\dim_\field\Hom{\Lambda}(FX,FY)$, where~$\Lambda$ is the cluster-tilted algebra~$\End{\cat}(T)$ and~$F$ is the equivalence of categories~$\cat(T,-):\cat/(\susp T) \to \MOD \Lambda$.

\noindent Theorem (see \cref{theo::relations-g-vecteurs}): \emph{The set~$L_\cat=\{\ell_X\,|\, X\in\ind(\cat)\setminus\add(\susp T)\}$ is a basis of the kernel of~$\b{g}$ and, for any~$x\in\ker(\b{g})$, we have:}
\[
x = \sum_{A \in \ind(\cat) \ssm \add(\susp T)} \frac{\langle x, A \rangle}{\langle \ell_A, A \rangle} \ell_A.
\]

\bigskip

\noindent {\bf Extriangulated categories}

In \cref{sec:extricats} we prove a more general version of \cref{theo::relations-g-vecteurs} that not only applies to cluster fans, but also to non-kissing fans.
Once more, our main motivation is to prove that the type cone of the non-kissing fan of a brick and 2-acyclic gentle algebra is simplicial, but the results are of independent interest, and might be useful for studying the type cones of other families of simplicial fans.

Let~$\cat$ be an extriangulated category (\cref{sect::recollections extricats}) with a fixed projective object~$T$ such that the morphism~$T\to 0$ is an inflation and, for each~$X\in\cat$, there is an extriangle~$T_1^X\infl T_0^X\defl X\overset{\delta_X}{\dashrightarrow}$.
Let~$\susp T$ be the cone of the inflation~$T\infl 0$.

\noindent \cref{prop::KRextricat}: \emph{The functor~$\cat(T,-)$ induces an equivalence~$F:\cat/(\susp T) \to \MOD \End{\cat}(T)$.}

@@

\noindent \cref{prop::index iso extricat}: \emph{The index induces an isomorphism of abelian groups from~$\kzero{\cat}$ to~$\kzero{\add T}$.}

@@

\noindent \cref{theorem::extricats}: \emph{The extriangulated category~$\cat$ has only finitely many ismorphism classes of indecomposable objects if and only if the set~$L_\cat$ generates the kernel of the canonical projection~$\b{g}:\Ksp(\cat)\to\kzero{\cat}$.
In that case~$L_\cat$ is a basis of the kernel of~$\b{g}$ and any~$x\in\ker(\b{g})$ decomposes as}
\[
 x = \sum_{A \in \ind(\cat) \ssm \add(\susp T)} \frac{\langle x, A \rangle}{\langle \ell_A, A \rangle} \ell_A.
\]

\bigskip

\noindent {\bf Schematic diagram of logical dependencies:}

@@

% Schematic diagram:
% \yann{Vincent, Arnau, je suis certain que vous feriez bien mieux que moi. Puis-je vous demander de le représenter ?}
% \arnau{As tu une idée en tête de comment le répresenter? Je trouve que ce que t'as écrit est très clair et je ne vois pas comment faire une schema qui le ameillore}
% 
% I.1 explains the type cone approach to the problem
% 
% This strategy is used in I.2.2 for cluster algebras and in I.2.3 for gentle algebras.
% 
% II.3 cluster categories (2-Calabi--Yau triangulated categories) is used in 1.2.2 for proving U.E.R.P. and simpliciality of the type cone.
% 
% II.4 generalises II.3, is used in I.2.3 for proving the U.E.R.P. and offers a different proof of simpliciality of the type cone .

@@

The following statement is proved in \cref{coro::number of positive meshes}.

@@

\begin{proof}
Points~(i), (iii) and~(iv) were shown in~\cite[Prop.~2.33, Thm.~4.17 \& Prop~4.16]{PaluPilaudPlamondon-nonkissing}. We postpone the proof of point~(ii) until \cref{coro::exchangeablePairsNKC}.
\end{proof}

@@

The following statement describes the type cone of the non-kissing fan of a brick and $2$-acyclic gentle quiver.
We provide here an elementary combinatorial proof, a more general representation theoretic perspective is discussed in \cref{theo::brick-algebra-condition}.

@@

Our next statement follows from the end of the previous proof.
See also \cref{coro::simplicialTypeConeNKCalgebraic}.

@@

In~\cite{Auslander1984}, M.~Auslander showed that the relations in the Grothendieck group of an Artin algebra are generated by the ones given by almost-split sequences precisely when the algebra is of finite representation type. Inspired by this result, we will show that a similar statement holds (see \cref{theo::relations-g-vecteurs}) for the~$\b{g}$-vectors (or indices) of objects in a~$2$-Calabi--Yau triangulated category.

@@

\subsection{Setting}
\label{sect::setting}

@@

\subsection{Statement of the theorem}
\label{subsect::Statement cluster cats}

@@

\begin{proposition}[\cite{BuanMarshReiten, KellerReiten}]\label{prop::functor-F}

@@

\begin{definition}
\label{defi::grothendieck-group}
\begin{enumerate}
\item Let~$\Ksp(\cat)$ be the \defn{split Grothendieck group} of~$\cat$, that is, the free abelian group generated by symbols~$[X]$, where~$[X]$ denotes the isomorphism class of~$X$ in~$\cat$, modulo the following relations: for any objects~$Y$ and~$Z$, we let~$[Y\oplus Z] = [Y] + [Z]$.

@@

In particular,~$\Ksp(\cat)$ is isomorphic to a free abelian group over the set~$\ind(\cat)$. Considering the modified group~$K_0(\cat ; T)$ instead of the usual Grothendieck group is motivated by \cref{sect::applications-g-vectors}, where we study relations between~$\b{g}$-vectors.

\begin{definition}
\label{defi::bilinear form}
For any two objects~$X$ and~$Y$ of~$\cat$, define
\[
\langle X, Y \rangle \eqdef \dim_{\field} \Hom{\Lambda}(FX,FY).
\]
This defines a bilinear form
\[
\langle -,-\rangle : \Ksp(\cat) \times \Ksp(\cat) \xrightarrow{} \Z.
\]
\end{definition}

\begin{notation}
\label{notation::ell_X}
\begin{enumerate}
\item For any indecomposable object~$X$ of~$\cat$, let
\[
X \to E \to \susp^{-1} X \to \susp X
\]
be an almost split triangle (unique up to isomorphism). We let
\[
\ell_X \eqdef [X] + [\susp^{-1} X]  - [E] \in \Ksp(\cat).
\]

@@

\begin{theorem}
\label{theo::relations-g-vecteurs}
Let~$\cat$ be a category satisying the hypotheses of \cref{sect::setting}. Then~$\cat$ has only finitely many isomorphism classes of indecomposable objects if and only if the set
\[
L \eqdef \bigset{\ell_X}{X \in \ind(\cat) \ssm \add(\susp T)}
\]
generates the kernel of~$\b{g}:\Ksp(\cat) \to K_0(\cat ; T)$. 
In this case, the set~$L$ is a basis of the kernel of~$\b{g}$, and for any $x \in \ker(\b{g})$, we have that
\[
x= \sum_{A \in \ind(\cat) \ssm \add(\susp T)} \frac{\langle x, A \rangle}{\langle \ell_A, A \rangle} \ell_A.
\]
\end{theorem}

\begin{corollary}
\label{corollary::meshes positively generate cluster cats}
Assume that~$\ind(\cat)$ is finite. Let~$X\xrightarrow{} E \xrightarrow{} Y \xrightarrow{h} \susp X$ be a triangle with~$h \in (\susp T)$. Then the element~$x=[X]+[Y]-[E]$ of the kernel of~$\b{g}$ is a non-negative linear combination of the~$\ell_A$, with~$A \in \ind(\cat)\ssm \add(\susp T)$.
\end{corollary}

\begin{proof}
We know that~$x=\sum_{A \in \ind(\cat) \ssm \add(\susp T)} \frac{\langle x, [A] \rangle}{\langle \ell_A, [A] \rangle} \ell_A$; since~$\langle \ell_A, [A] \rangle$ is positive by \cref{lemm::bilinear-form}, we only need to show that each~$\langle x, [A] \rangle$ is non-negative. 
The functor~$F=\cat(T,-)$ induces an exact sequence
\[
FX \to FE \to FY \to 0,
\]
which in turn induces an exact sequence
\[
0\to \Hom{\Lambda}(FY, FA) \to \Hom{\Lambda}(FE, FA) \xrightarrow{f} \Hom{\Lambda}(FX, FA) \to \coker(f) \to 0.
\]
Therefore~$\langle x, [A] \rangle = \dim_{\field} \coker(f) \ge 0$.
\end{proof}

@@

\subsection{Proof of \cref{theo::relations-g-vecteurs}}

Before we can prove \cref{theo::relations-g-vecteurs}, we need to recall the following results and definition from~\cite{DehyKeller,Palu}.

@@

The notion of index is very close to the definition of the map~$\b{g}:\Ksp(\cat)\to K_0(\cat ; T)$ of \cref{defi::grothendieck-group}. The link is given by the following result.

@@

\begin{corollary}
\label{coro::grothendieck-g-vectors}
There is an isomorphism~$\phi:K_0(\cat ; T) \to \Ksp \big( \add(T) \big)$ such that~$\ind_T = \phi\circ \b{g}$. In particular,~$K_0(\cat ; T)$ is a free abelian group generated by the~$[T_i]$.
\end{corollary}

We can now begin the proof of \cref{theo::relations-g-vecteurs}

\begin{lemma}
\label{lemm::bilinear-form-suspention-T}
If~$X$ or~$Y$ lie in~$\add(\susp T)$, then~$\langle X, Y \rangle = 0$.
\end{lemma}

@@

\begin{lemma}
\label{lemm::almost-split}
Let~$X\xrightarrow{} Y \xrightarrow{} \susp^{-1} X \xrightarrow{h} \susp X$ be an almost-split triangle. Then~$X\notin \add(\susp T)$ if and only if~$h \in (\susp T)$.
\end{lemma}

@@

\begin{lemma}
\label{lemm::bilinear-form}
Let~$A$ and~$B$ be two indecomposable objects of~$\cat$.
\begin{enumerate}
\item If~$A \notin \add(\susp T)$, then
\[
\langle \ell_A, B \rangle = 
\begin{cases}
0 & \textrm{if~$A\ncong B$;} \\
\dim_{\field}\field_A & \textrm{if~$A\cong B$.}
\end{cases}
\]

@@

\subsection{Application to~$\b{g}$-vectors of cluster algebras of finite type}
\label{sect::applications-g-vectors}

>>>>

\section*{Structure of the paper: main results and logical dependencies}

@@

\para{Type cone approach}
(\cref{sec:typeCone})

@@

\para{Cluster algebras of finite type}
(\cref{subsec:typeConeCA})

\noindent
We first consider the~$\b{g}$-vector fan~$\Fan(\B_\circ)$ of a skew-symmetrizable cluster algebra of finite type, with respect to any initial exchange matrix of finite type~$\B_\circ$. We provide the following polytopal realizations of~$\Fan(\B_\circ)$:
% We consider the set~$\meshes$ of all pairs~$\{x,x'\}$ related by non-initial mesh mutations (\cref{def:meshMutation}), and we denote by~$\principalVariables$ the set of cluster variables.
% Let~$\{x,x'\}\in\meshes$, and let~$X,X'$ be two clusters such that~$X\setminus\{x\}=X'\setminus\{x'\}$.
% For any~$y\in\principalVariables$, define~$\alpha_{x,x',y}$ to be~$|b_{xy}|$ if~$y\in X\cap X'$ and 0 otherwise.
% By the unique exchange relation property, the~$\alpha_{x,x',y}$'s do not depend on the specific choice of the clusters~$X,X'$.

\newtheorem*{thm:allPolytopalRealizationsCA}{\cref{thm:allPolytopalRealizationsCA}}
\begin{thm:allPolytopalRealizationsCA}
For any~$\b{\ell} \in \R_{>0}^{\meshes}$, the polytope
\[
R_{\b{\ell}}(\B_\circ) \eqdef \Bigset{\b{z} \in \R^{\principalVariables}}{\b{z} \ge 0 \text{ and } \b{z}_x + \b{z}_{x'} - \sum_{y \in \principalVariables} \alpha_{x,x',y} \, \b{z}_{y} = \b{\ell}_{\{x,x'\}} \text{ for all } \{x,x'\} \in \meshes}
\]
is a generalized associahedron, whose normal fan is the cluster fan~$\gvectorFan[\B_\circ]$.
Moreover, the polytopes~$R_\b{\ell}(\B_\circ)$ for~$\b{\ell} \in \R_{>0}^{\meshes}$ describe all polytopal realizations of~$\gvectorFan[\B_\circ]$.
\end{thm:allPolytopalRealizationsCA}

@@

Some crucial pieces of the proof of~\cref{thm:allPolytopalRealizationsCA} use cluster categories (2-Calabi--Yau triangulated categories) and are deferred to~\cref{sec:clusterCategories}. Namely, a reformulation of~\cite[Thm.~7.5]{BuanMarshReinekeReitenTodorov}, proved in~\cref{sec:clusterCategories}, gives the first step towards applying our type cone strategy.

\newtheorem*{prop:uniqueExchangePropertyCA}{\cref{prop:uniqueExchangePropertyCA}}
\begin{prop:uniqueExchangePropertyCA}
The cluster fan~$\Fan(\B_\circ)$ has the unique exchange relation property.
\end{prop:uniqueExchangePropertyCA}

@@

\newtheorem*{coro:simplicialTypeConeCA}{\cref{coro:simplicialTypeConeCA}}
\begin{coro:simplicialTypeConeCA}
The type cone~$\typeCone(\Fan(\B_\circ))$ is simplicial.
\end{coro:simplicialTypeConeCA}

\para{Gentle algebras, non-kissing complex, generalized accordiohedra}
(\cref{subsec:typeConeNKC})

\noindent
The same strategy is applied to a class of finite dimensional algebras, called gentle algebras, whose representations are combinatorially well understood.
We identify two conditions, called brick and 2-acyclicity (\cref{def:brick2acyclic}), that enable us to make use of the type cone approach.
We note that the~$\tau$-tilting theory of gentle algebras is the algebraic counterpart of the combinatorics of accordions on surfaces~\cite{PaluPilaudPlamondon-nonkissing,BrustleDouvilleMousavandThomasYildirim,PaluPilaudPlamondon-surfaces}.

Let~$\quiver = (Q,I)$ be a gentle bound quiver, and let~$\Fan(\quiver)$ be the fan of its~$\b{g}$-vectors (\cref{subsubsec:nonkissingComplex}), also called its non-kissing fan.
We prove \cref{prop:exchangeablePairsNKC} (ii) in \cref{sec:Kbproj}, by making use of representation theory, and state here the following consequence.

\newtheorem*{corollary: uerp gentle}{\cref{corollary: uerp gentle}}
\begin{corollary: uerp gentle}
The non-kissing fan~$\Fan(\quiver)$ has the unique exchange relation property.
\end{corollary: uerp gentle}

\cref{subsubsec:simplicialTypeConeNKC} is dedicated to giving a purely combinatorial proof of the following statement.

\newtheorem*{coro:simplicialTypeConeNKC}{\cref{coro:simplicialTypeConeNKC}}
\begin{coro:simplicialTypeConeNKC}
Assume that the gentle bound quiver~$\quiver$ is brick and 2-acyclic.
Then the type cone~$\typeCone(\Fan(\quiver))$ is simplicial.
\end{coro:simplicialTypeConeNKC}

An alternative algebraic proof, using the main result of \cref{sec:extricats}, is given in \cref{sec:Kbproj}.

Combining these pieces, we make use of our type cone approach to deduce the following statement.

\newtheorem*{coro:allPolytopalRealizationsNKC}{\cref{coro:allPolytopalRealizationsNKC}}
\begin{coro:allPolytopalRealizationsNKC}
For any brick and $2$-acyclic gentle quiver~$\quiver$ and any~$\b{\ell} \in \R_{>0}^{\strings}$, the polytope
\[
R_\b{\ell}(\quiver) \eqdef \set{\b{z} \in \R^{\walks}}{\begin{array}{l} \b{z} \ge 0 \qquad\text{and}\qquad \b{z}_\omega = 0 \text{ for any improper walk } \omega \\ \b{z}_{\cc{\sigma}} + \b{z}_{\hh{\sigma}} - \b{z}_{\hc{\sigma}} - \b{z}_{\ch{\sigma}} = \b{\ell}_\sigma \text{ for all } \sigma \in \strings\end{array}}
\]
is a realization of the non-kissing fan~$\gvectorFan[\quiver]$.
Moreover, the polytopes~$R_\b{\ell}(\quiver)$ for~$\b{\ell} \in \R_{>0}^{\strings}$ describe all polytopal realizations of~$\gvectorFan[\quiver]$.
\end{coro:allPolytopalRealizationsNKC}

@@

\para{Graphical nested complexes}
(\cref{subsec:typeConeGA})

@@

\newtheorem*{cor:typeconenestedfan}{\cref{cor:typeconenestedfan}}
\begin{cor:typeconenestedfan}
Consider a graph~$\graphG$ on~$\ground$ with tubes~$\tubes$. 
Then, for a height vector~$\b{h} \in \R^{\tubes}$, the nested fan~$\gvectorFan[\graphG]$ is the normal fan of the graph associahedron
\[
\set{\b{x} \in \R^\ground}{\dotprod{\gvector{\tube}}{\b{x}} \le \b{h}_{\tube} \text{ for any tube } \tube \in \tubes}
\]
if and only if~$\b{h}_{\varnothing} = \b{h}_{\graphG} = 0$ and
 $\b{h}_{\tube[s] \ssm \{v'\}} + \b{h}_{\tube[s] \ssm \{v\}} > \b{h}_{\tube[s]} + \b{h}_{\tube[s] \ssm \{v, v'\}}$ for any tube~$\tube[s] \in \tubes$ and distinct non-disconnecting vertices~$v,v'$ of~$\tube[s]$.
\end{cor:typeconenestedfan}

\para{Cluster categories}
(\cref{sec:clusterCategories})

@@

Let~$\cat$ be a cluster category with finitely many isomorphism classes of indecomposable objects (see \cref{sec:setting} for the precise and more general setting in which the theorem is proven).
Let~$\Ksp(\cat)$ be the split Grothendieck group of~$\cat$.
Fix a cluster-tilting object~$T\in\cat$, and let~$K_0(\cat ; T)$ be the quotient of~$\Ksp(\cat)$ by the relations~$[X]+[Z]-[Y]$ for all triangles
$
X\xrightarrow{} Y \xrightarrow{} Z \xrightarrow{h} \susp X
$
with~$h \in (\susp T)$.
Denote by~$\b{g}:\Ksp(\cat) \to K_0(\cat ; T)$ the canonical projection.
For any indecomposable~$X\in\cat$, write~$\ell_X = [X]+[\susp^{-1}X]-[E]$, where~$E$ is the middle term of an almost split triangle starting at~$X$.
For any objects~$X,Y\in\cat$, we write~$\langle X, Y \rangle$ for~$\dim_\field\Hom{\Lambda}(FX,FY)$, where~$\Lambda$ is the cluster-tilted algebra~$\End{\cat}(T)$ and~$F$ is the equivalence of categories~$\cat(T,-):\cat/(\susp T) \to \MOD \Lambda$.

\newtheorem*{thm:relations-g-vecteurs}{\cref{thm:relations-g-vecteurs}}
\begin{thm:relations-g-vecteurs}
The set~$L_\cat \eqdef \set{\ell_X}{X \in \ind(\cat) \setminus \add(\susp T)}$ is a basis of the kernel of~$\b{g}$ and, for any~$x \in \ker(\b{g})$, we have:
\[
x = \sum_{A \in \ind(\cat) \ssm \add(\susp T)} \frac{\langle x, A \rangle}{\langle \ell_A, A \rangle} \ell_A.
\]
\end{thm:relations-g-vecteurs}

\para{Extriangulated categories}
(\cref{sec:extricats})

In \cref{sec:extricats} we prove a more general version of \cref{thm:relations-g-vecteurs} that not only applies to cluster fans, but also to non-kissing fans.
Once more, our main motivation is to prove that the type cone of the non-kissing fan of a brick and 2-acyclic gentle algebra is simplicial, but the results are of independent interest, and might be useful for studying the type cones of other families of simplicial fans.

Let~$\cat$ be an extriangulated category (\cref{sec:recollections extricats}) with a fixed projective object~$T$ such that the morphism~$T\to 0$ is an inflation and, for each~$X\in\cat$, there is an extriangle~$T_1^X\infl T_0^X\defl X\overset{\delta_X}{\dashrightarrow}$.
Let~$\susp T$ be the cone of the inflation~$T\infl 0$.

\newtheorem*{prop:KRextricat}{\cref{prop:KRextricat}}
\begin{prop:KRextricat}
The functor~$\cat(T,-)$ induces an equivalence~$F:\cat/(\susp T) \to \MOD \End{\cat}(T)$.
\end{prop:KRextricat}

@@

\newtheorem*{prop:index iso extricat}{\cref{prop:index iso extricat}}
\begin{prop:index iso extricat}
The index induces an isomorphism of abelian groups from~$\kzero{\cat}$ to~$\kzero{\add T}$.
\end{prop:index iso extricat}

@@

\newtheorem*{thm:extricats}{\cref{thm:extricats}}
\begin{thm:extricats}
The extriangulated category~$\cat$ has only finitely many ismorphism classes of indecomposable objects if and only if the set~$L_\cat$ generates the kernel of the canonical projection~$\b{g}:\Ksp(\cat)\to\kzero{\cat}$.
In that case~$L_\cat$ is a basis of the kernel of~$\b{g}$ and any~$x\in\ker(\b{g})$ decomposes as
\[
 x = \sum_{A \in \ind(\cat) \ssm \add(\susp T)} \frac{\langle x, A \rangle}{\langle \ell_A, A \rangle} \ell_A.
\]
\end{thm:extricats}

\para{Schematic diagram of logical dependencies}

@@

@@

The following statement is proved in \cref{coro:number of positive meshes}.

@@

\begin{proof}
Points~(i), (iii) and~(iv) were shown in~\cite[Prop.~2.33, Thm.~4.17 \& Prop~4.16]{PaluPilaudPlamondon-nonkissing}. We postpone the proof of point~(ii) until \cref{coro:exchangeablePairsNKC}.
\end{proof}

@@

The following statement describes the type cone of the non-kissing fan of a brick and $2$-acyclic gentle quiver.
We provide here an elementary combinatorial proof, a more general representation theoretic perspective is discussed in \cref{thm:brick-algebra-condition}.

@@

Our next statement follows from the end of the previous proof.
See also \cref{coro:simplicialTypeConeNKCalgebraic}.

@@

In~\cite{Auslander1984}, M.~Auslander showed that the relations in the Grothendieck group of an Artin algebra are generated by the ones given by almost-split sequences precisely when the algebra is of finite representation type. Inspired by this result, we will show that a similar statement holds (see \cref{thm:relations-g-vecteurs}) for the~$\b{g}$-vectors (or indices) of objects in a~$2$-Calabi--Yau triangulated category.

@@

\subsection{Setting}
\label{sec:setting}

@@

\subsection{Statement of the theorem}
\label{subsec:Statement cluster cats}

@@

\begin{proposition}[\cite{BuanMarshReiten, KellerReiten}]\label{prop:functor-F}

@@

\begin{definition}
\label{def:grothendieck-group}
\begin{enumerate}
\item Let~$\Ksp(\cat)$ be the \defn{split Grothendieck group} of~$\cat$, that is, the free abelian group generated by symbols~$[X]$, where~$[X]$ denotes the isomorphism class of~$X$ in~$\cat$, modulo the following relations: for any objects~$Y$ and~$Z$, we let~$[Y\oplus Z] = [Y] + [Z]$.

@@

In particular,~$\Ksp(\cat)$ is isomorphic to a free abelian group over the set~$\ind(\cat)$. Considering the modified group~$K_0(\cat ; T)$ instead of the usual Grothendieck group is motivated by \cref{sec:applications-g-vectors}, where we study relations between~$\b{g}$-vectors.

\begin{definition}
\label{def:bilinear form}
For any two objects~$X$ and~$Y$ of~$\cat$, define
\[
\langle X, Y \rangle \eqdef \dim_{\field} \Hom{\Lambda}(FX,FY).
\]
This defines a bilinear form
\[
\langle -,-\rangle : \Ksp(\cat) \times \Ksp(\cat) \xrightarrow{} \Z.
\]
\end{definition}

\begin{notation}
\label{notation:ell_X}
\begin{enumerate}
\item For any indecomposable object~$X$ of~$\cat$, let
\[
X \to E \to \susp^{-1} X \to \susp X
\]
be an almost split triangle (unique up to isomorphism). We let
\[
\ell_X \eqdef [X] + [\susp^{-1} X]  - [E] \in \Ksp(\cat).
\]

@@

\begin{theorem}
\label{thm:relations-g-vecteurs}
Let~$\cat$ be a category satisying the hypotheses of \cref{sec:setting}. Then~$\cat$ has only finitely many isomorphism classes of indecomposable objects if and only if the set
\[
L \eqdef \bigset{\ell_X}{X \in \ind(\cat) \ssm \add(\susp T)}
\]
generates the kernel of~$\b{g}:\Ksp(\cat) \to K_0(\cat ; T)$. 
In this case, the set~$L$ is a basis of the kernel of~$\b{g}$, and for any $x \in \ker(\b{g})$, we have that
\[
x= \sum_{A \in \ind(\cat) \ssm \add(\susp T)} \frac{\langle x, A \rangle}{\langle \ell_A, A \rangle} \ell_A.
\]
\end{theorem}

\begin{corollary}
\label{coro:meshes positively generate cluster cats}
Assume that~$\ind(\cat)$ is finite. Let~$X\xrightarrow{} E \xrightarrow{} Y \xrightarrow{h} \susp X$ be a triangle with~$h \in (\susp T)$. Then the element~$x=[X]+[Y]-[E]$ of the kernel of~$\b{g}$ is a non-negative linear combination of the~$\ell_A$, with~$A \in \ind(\cat)\ssm \add(\susp T)$.
\end{corollary}

\begin{proof}
We know that~$x=\sum_{A \in \ind(\cat) \ssm \add(\susp T)} \frac{\langle x, [A] \rangle}{\langle \ell_A, [A] \rangle} \ell_A$; since~$\langle \ell_A, [A] \rangle$ is positive by \cref{lem:bilinear-form}, we only need to show that each~$\langle x, [A] \rangle$ is non-negative. 
The functor~$F=\cat(T,-)$ induces an exact sequence
\[
FX \to FE \to FY \to 0,
\]
which in turn induces an exact sequence
\[
0\to \Hom{\Lambda}(FY, FA) \to \Hom{\Lambda}(FE, FA) \xrightarrow{f} \Hom{\Lambda}(FX, FA) \to \coker(f) \to 0.
\]
Therefore~$\langle x, [A] \rangle = \dim_{\field} \coker(f) \ge 0$.
\end{proof}

@@

\subsection{Proof of \cref{thm:relations-g-vecteurs}}

Before we can prove \cref{thm:relations-g-vecteurs}, we need to recall the following results and definition from~\cite{DehyKeller,Palu}.

@@

The notion of index is very close to the definition of the map~$\b{g}:\Ksp(\cat)\to K_0(\cat ; T)$ of \cref{def:grothendieck-group}. The link is given by the following result.

@@

\begin{corollary}
\label{coro:grothendieck-g-vectors}
There is an isomorphism~$\phi:K_0(\cat ; T) \to \Ksp \big( \add(T) \big)$ such that~$\ind_T = \phi\circ \b{g}$. In particular,~$K_0(\cat ; T)$ is a free abelian group generated by the~$[T_i]$.
\end{corollary}

We can now begin the proof of \cref{thm:relations-g-vecteurs}

\begin{lemma}
\label{lem:bilinear-form-suspention-T}
If~$X$ or~$Y$ lie in~$\add(\susp T)$, then~$\langle X, Y \rangle = 0$.
\end{lemma}

@@

\begin{lemma}
\label{lem:almost-split}
Let~$X\xrightarrow{} Y \xrightarrow{} \susp^{-1} X \xrightarrow{h} \susp X$ be an almost-split triangle. Then~$X\notin \add(\susp T)$ if and only if~$h \in (\susp T)$.
\end{lemma}

@@

\begin{lemma}
\label{lem:bilinear-form}
Let~$A$ and~$B$ be two indecomposable objects of~$\cat$.
\begin{enumerate}
\item If~$A \notin \add(\susp T)$, then
\[
\langle \ell_A, B \rangle = 
\begin{cases}
0 & \textrm{if~$A\ncong B$;} \\
\dim_{\field}\field_A & \textrm{if~$A\cong B$.}
\end{cases}
\]

@@

\subsection{Application to~$\b{g}$-vectors of cluster algebras of finite type}
\label{sec:applications-g-vectors}
